Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-10-20 Manuscript/manuscript.tex
Updates

\subsubsection{Simulations of POPC bilayers with aqueous ions}
Simulations of POPC bilayer in pure water or in varying salt concentrations
contained 128 POPC molecules and approximately 50~water molecules per each lipid
in the periodic orthorhombic simulation box. The water was described using 
SPC/E \cite{Berendsen1987} model. This model has the optimal dielectric constant
to be used with ECC modified molecules \todo{Citation or discussion.} and has been 
was also used in the parametrization of ECC
ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.
The results with other water models are available in SI.
%As a default, water molecules were described by the OPC3 force field~\cite{Izadi16} 
%which is currently the most accurate three site rigid water model
%In order to test transferability of our newly developed ECC-lipids model, 
%we also performed several additional simulations with 
%OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987}, TIP3p-FB and TIP4p-FB~\cite{Wang2014},
%and TIP4p/2005 \cite{Abascal2005} water models
%\todo{The normal TIP3P was tested as well, right?}
%presented in Supporting Information (SI). 
The ECC-ion models \todo{Specify the exact version.} were used for Sodium, Calcium and Chloride
ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
Simulation with the Lipid14 model was also ran with the ion model by Dang et al.~\cite{smith94,chang1999,dang2006}.
Simulation data for the Lipid14 model with \AA{}qvist~\cite{aqvist90} ion model
was taken directly from \cite{lipid14POPC0mMNaClfiles}. \todo{Which water model was used in these simulations?} 
MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).  
The simulation settings used in this work are 
summarized in Table \ref{tbl:mdpar}. 
%and they are based on previously used settings in \cite{catte16} available at \cite{lipid14POPC0mMNaClfiles}.
%\todo{As far as remember, I used there Langevin dynamics instead of thermostated MD, because
%this is done in Amber by default. If this is correct, the information in the
%table do not match with this sentence. Based on semi-extensive testing I made few years
%ago this do not change anything. Anyway, this should be reported consistently.}
Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.

\subsubsection{Simulations of POPC bilayers with cationic surfactants}
An automated topology builder \cite{malde11} was first used to create the structure of
dihexadecyldimethylammonium.
The AmberTools program \cite{amber} was then used to generate the Amber-type force field
parameters. The parameters were converted to the Gromacs format by using
the acpype tool \cite{acpype}. The partial charges were then manually modified
to approximately correspond to their equivalent segments in Lipid14 \cite{dickson14}.
The parameters together with the simulation data are available
at~\cite{POPClipid14T313K,POPClipid1410perCATsurfT313K,POPClipid1420perCATsurfT313K,POPClipid1430perCATsurfT313K,POPClipid1442perCATsurfT313K,POPClipid1450perCATsurfT313K}.
The surfactants were randomly placed among the lipids to form bilayer structures with
mole fractions of 10\%, 20\%, 30\%, 42\%, or 50\% of surfactant in the POPC bilayer.
All systems contained 50 POPC molecules per leaflet, 6340 TIP3P water molecules and
6, 14, 21, 35, or 50 surfactants per leaflet. 
Chloride counter ions were used
in simulations, because parameters for bromide, the counterion in the experimental data \cite{scherer89},
were not available in the standard Gromacs files for Amber force field.
\todo{JOE: We cannot present an intricate cool new model for a lipid and claim at the same time that we can't simulate bromide. }
The Lipid14 model was used for POPC.
The first 20~ns of the total simulation time of 200~ns
was considered as an equlibration time and was omitted from the analysis.
A reasonable lipid neighbor exchange occured during the simulation.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=7.5cm]{../Fig/POPCstructure.eps}
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.pdf}
  \includegraphics[height=8.4cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
  \caption{\label{simVSexpNOions}
    Top: The chemical structure of POPC and the labeling of the carbon segments.
    Middle: Order parameters of POPC head group, glycerol backbone and acyl chains 
    from simulations with the Lipid14 \cite{dickson14} and the ECC-lipid models
    compared with the experimental values from \cite{ferreira13}.
    The size of the points for the head group order parameters correspond the
    of the error bars ($\pm 0.02$ for experiments \cite{botan15,ollila16} and ?? for simulations).
    The size of the points for acyl chains are decreased by a factor of 3 to improve the clarity of the plot.
    Bottom: X-ray scattering form factors from simulations with the Lipid14 \cite{dickson14} and
    the ECC-lipid models compared with experiments~\cite{Kucerka2011}.
  } 
  \todo{Increase the height of the order parameter figure} \\
  \todo{sn-1 and sn-2 order parameters should be somehow labelled in the order parameter figure. Maybe empty and filled points?} \\
  \todo{Add size of the error bars in simulations to the caption.} \\
  \todo{Would it be possible to increase the size of points for acyl chain order parameters only in z-direction such that
    it would correspond the error bars?} \\
  \todo{x-axis in form factor plot from 0 to 0.6, where experimental data ends. y-axis from 0 to $\sim$2.5 to remove the empty space.} \\
\end{figure}

\begin{table}
  \caption{Area per lipid (APL) values of POPC with no ions from the Lipid14 simulation
    ran in this work and from the literature, from ECC-lipid model and from experiments.\label{tab:apls} }
  \begin{tabular}{l|c c}
    model          & APL (\AA$^2$)   & Temperature [K] \\
    \hline
    Lipid14                   & 65.1$\pm$ 0.6  &  300 \\
    Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
    \hline
    ECC-lipids                & 62.2$\pm$ 0.6  &  300       \\
    \hline
    experiment \cite{kucerka11} & 64.3  &  303    \\
    \hline
  \end{tabular}
\end{table}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl.pdf}
  \caption{\label{fig:delta_ordPar_CaCl}
    Changes of head group order parameters of POPC bilayer as a function of CaCl$_2$ concentrations
    are shown from simulations with different force fields together with experimental data 
    (DPPC \cite{akutsu81} and POPC \cite{altenbach84}). 
    Ion concentrations in bulk water are shown in x-axis. 
    Values from simulations are calculated from the of cation number density $C_{np}$
    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
    Simulation data with Lipid14 and \AA{}qvist ion parameters is taken directly from Ref. \cite{catte16}.
  }
  \todo{I think that we must show the NaCl results in the main paper. I think that the best would
    be to use similar two column format, which we had before. This is used also in NMRlipids II and makes the
    comparison easier} \\
  \todo{The DPPC and POPC should be shown and properly labelled as we had previously.} \\
  \todo{The labeling should be consistent with previous figures, i.e., experiments with diamonds.} \\
  \todo{Start x-axis scale from -1 to make the point in zero fully visible.} \\
  \todo{Start y-axis scale of the two top figures from -0.13 to remove the empty space.} \\
  \todo{y-axis scale of the bottom figure from 59 to 78 to remove the empty space.} \\
  \todo{Empty space between figures and from the right column could be reduced.} \\
  \todo{Experimental values from \cite{akutsu81} to be put in the $g_3$ figure:
    the value of $g_3$ order parameter of DPPC was -0.214 
    in the absence of ions [average of two closely spaced splittings] and -0.211 in the presence
    of 0.35 M CaCl$_2$ (at 59 Celcius). No effect of ions could be detected on DPPC bilayers labeled
    at the C-2 segments of both fatty acyl chains.''
  }
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/OrdPars-A-B_L14-ECCL17_q80_sig89_NaCl.pdf}
  \caption{\label{fig:delta_ordPar_NaCl}
    Changes of head group order parameters $\alpha$ and $\beta$ of POPC bilayer as a function of NaCl concentrations
    are shown from simulations with different force fields together with experimental data \cite{akutsu81}. 
    Ion concentrations in bulk water are shown in x-axis. 
    Values from simulations are calculated from the of cation number density $C_{np}$
    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
    Simulation data with Lipid14 and \AA{}qvist ion parameters is taken directly from Ref. \cite{catte16}.
  }
\end{figure}

@@

Since significant artefacts are reported in simulations with \AA{}qvist ions is water \cite{??},
we also simulated the Lipid14 model with the ion models by Dang et al.~\cite{smith94,chang1999,dang2006} and
ECC-ions~\cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014} having
more realistic bulk behaviour. Instead of the improvement in the binding behaviour,
we observed overbinding also of Na$^+$ with these ions models, while the results
with Ca$^{2+}$ were barely affected,
%for the CaCl$_2$ interactions with a POPC bilayer described by the Lipid14 model,
as seen in Figs. \ref{fig:delta_ordPar_CaCl} and S??.%\ref{fig:ordPars_waterModels} (in SI), respectively.
\todo{Add OP-response of Lipid14+ECC-ions plot in SI}.
The results are in line with the previous work \cite{catte16},
suggesting that the improvements in the lipid parameters are
required to correctly describe the divalent cation binding to PC lipid bilayers.

%To improve the cation binding behaviour to POPC lipid bilayer we implicitly included the electronic
%polarization to the Lipid14 model by using the ECC approach as described in section \ref{section:ecc}.
The results from the simulations with the ECC-lipid and the ECC-ion models \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}
exhibit an improved behaviour of cation binding to a POPC bilayer in Fig.~\ref{fig:delta_ordPar_CaCl},
showing a good agreement with experiments in the changes of 
the lipid headgroup order parameters as a function of NaCl and CaCl$_2$ concentrations.
Since also the headgroup order parameter
response the to the bound positive charge the ECC-lipid model
was in good agreement with experiments in section \ref{section:boundCHARGE},
we conclude that the model correctly reproduces the binding affinity of
Na$^{+}$ and Ca$^{2+}$ ions to POPC lipid bilayer.
%In conclusion, the results suggest that 
%the ECC-lipid model together with the ECC-ion parameters gives a realistic
%description for the binding affinities of Na$^+$ and Ca$^{2+}$ ions to POPC lipid bilayer. 
Furthermore, the overestimated lipid headgroup order parameter changes of POPC in the Lipid14
model as a function of CaCl$_2$ concentration arise from both, the overestimated
binding affinity and sensitivity of the headgroup tilt to the bound positive charge.
This probably applies also to the other lipid models considered in the previous study~\cite{catte16},
emphasizing the importance of the comparison of the lipid headgroup order parameter response
to the bound charge between simulations and experiments as done in section \ref{section:boundCHARGE}.

\todo{SAMULI: Maybe we should discuss the repeat distances and
  area per molecules measured at \cite{petrache06,pabst07,uhrikova08}
  }

@@

It was found in the original work~\cite{altenbach84} that 
a ternary complex binding model (i.e.~2~POPC:1~\ce{Ca^{2+}})
provides the best fit to experimental measurements of all considered models in that study. 
In such a model, there is a linear relationship between quantities 
$C_b$, mole fraction of bound \ce{Ca^{2+}} per POPC, and $\sqrt{C_b/C_I}$, 
where $C_I$ is the concentration of free cations at the plane of ion binding~\cite{altenbach84}.
The concentration $C_b$ was obtained from an extrapolation of linear relation 
between deuterium NMR measurements and atomic absorption spectroscopy for low concetrations of \ce{CaCl2}.
Such an extrapolation is valid as long as the mode of \ce{Ca^{2+}} binding 
remains constant throughout the extrapolation range. 
The concentration $C_I$ is determined by using the surface potential by using the Boltzmann equation.
However, Boltzmann theory yields inaccurate results
for divalent cations like \ce{Ca^{2+}}~\cite{Andelman1995}. 
An atomistic simulation, on the other hand, provides these quantities directly without severe assumptions.
\todo{Did you really calculate the $C_I$ from simulations without severe assumptions?
Note that this conctration at the plane of biding, which do not equal the concentration of free cations.}
Hence we hypothesise that the discrepancy between the results in the experiment~\cite{altenbach84} and 
our simulations likely lays in the fact that 
the assumptions and relations used for determinig concentrations $C_b$ and $C_I$ in the experiment~\cite{altenbach84}
gradually do not hold for higher concentrations of \ce{Ca^{2+}}.

@@

\begin{table}
  \caption{Area per lipid (APL) from different models of POPC with no ions\label{tab:apls_si} }
  \begin{tabular}{l|c c}
    model          & APL (\AA$^2$)   & Temperature [K] \\
    \hline
    Lipid14                   & 65.1$\pm$ 0.6  &  300 \\
    Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
    \hline
    ECC-lipids &        &  \\
    %Lipid14ecc0.80+sigma0.875 &        &  313    \\
    %($4.6\cdot 5.1 \, \mathrm{nm}^2$), 72 lipids patch, OPC3           & 63.2   &   313      \\
    ~ OPC3           & 62.2$\pm$ 0.6   &  300       \\
    ~ OPC3           & 64.2$\pm$ 0.6   &  313       \\
    ~ SPC/E          & 65.1$\pm$ 0.6   &  313       \\
    ~ OPC            & 64.4$\pm$ 0.6   &  313       \\
    ~ TIP4p/2005     & 66.8$\pm$ 0.6   &  313       \\
    %($9.2\cdot 10.2 \, \mathrm{nm}^2$), 288 lipids patch           & 65.5   &  313       \\ %% not done for this model with f_sigma=0.89
    %oMM small patch           & 63.65  &         \\
    %oMM 4xbig patch           & 63.7   &         \\
    \hline
    experiment   & 62.7  &  293    \\
    experiment \cite{jambeck12}\todoii{REF}{put original references, not Slipids param. paper.} & 64.3  &  303    \\
    experiment  & 67.3  &  323    \\
    experiment  & 68.1  &  333    \\
    %experiment POPE  & 56.6 &  303    \\
    \hline
  \end{tabular} \\
  \todo{Result with normal TIP3P missing?}
\end{table}

\begin{table}[btp]
  \caption{Simulation parameters}
  \label{tbl:mdpar}
  \begin{tabular}{ll}
    simulation property & parameter   \\
    \hline
    time-step           & 2~fs         \\
    equilibration time  & 100~ns  \\
    simulation time     & 200~ns  \\
    temperature         & 313~K       \\
    thermostat          & v-rescale  \cite{bussi07}   \\
    barostat            & Parrinello-Rahman, semi-isotropic \cite{parrinello81} \\
    long-range electrostatics & PME  \cite{darden93}  \\
    cut-off scheme      & Verlet \cite{Pall13}      \\
    Coulomb and VdW cut-off & 1.0~nm \\
    constraints         & LINCS, only hydrogen atoms \cite{hess97} \\
    constraints for water & SETTLE  \cite{miyamoto92} \\
    \hline
  \end{tabular}
\end{table}

>>>>

\subsubsection{Simulations of POPC bilayers with aqueous ions}
Simulations of POPC bilayer in pure water or in varying salt concentrations
contained 128 POPC molecules and approximately 50~SPC/E~water~\cite{Berendsen1987} molecules per each lipid
in the periodic orthorhombic simulation box. The chosen water model 
has the optimal dielectric constant
to be used with the ECC modified molecules \todo{Citation or discussion.} and it has been 
was also used in the previous parametrization of ECC
ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.
The results with other water models are available in SI.
%As a default, water molecules were described by the OPC3 force field~\cite{Izadi16} 
%which is currently the most accurate three site rigid water model
%In order to test transferability of our newly developed ECC-lipids model, 
%we also performed several additional simulations with 
%OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987}, TIP3p-FB and TIP4p-FB~\cite{Wang2014},
%and TIP4p/2005 \cite{Abascal2005} water models
%\todo{The normal TIP3P was tested as well, right?}
%presented in Supporting Information (SI). 
The ECC-ion models \todo{Specify the exact version.} were used for Sodium, Calcium and Chloride
ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
Simulation with the Lipid14 model was also ran with the ion model by
Dang et al.~\cite{smith94,chang1999,dang2006}\todo{Which water model was used in this simulation?}.
Simulation data for the Lipid14 model with \AA{}qvist~\cite{aqvist90} ion and TIP3P~\cite{jorgensen83}
water models are reported in \cite{catte16} and
were taken directly from \cite{lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC,lipid14POPC1000mMNaClfiles}. 
MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).  
The simulation settings used in this work are 
summarized in Table \ref{tbl:mdpar}. 
%and they are based on previously used settings in \cite{catte16} available at \cite{lipid14POPC0mMNaClfiles}.
%\todo{As far as remember, I used there Langevin dynamics instead of thermostated MD, because
%this is done in Amber by default. If this is correct, the information in the
%table do not match with this sentence. Based on semi-extensive testing I made few years
%ago this do not change anything. Anyway, this should be reported consistently.}
Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.

\subsubsection{Simulations of POPC bilayers with cationic surfactants}
An automated topology builder \cite{malde11} was first used to create the structure of
dihexadecyldimethylammonium.
The AmberTools program \cite{amber} was then used to generate the Amber-type force field
parameters. The parameters were converted to the Gromacs format by using
the acpype tool \cite{acpype}. The partial charges were then manually modified
to approximately correspond to their equivalent segments in Lipid14 \cite{dickson14}.
The parameters together with the simulation data are available
at~\cite{POPClipid14T313K,POPClipid1410perCATsurfT313K,POPClipid1420perCATsurfT313K,POPClipid1430perCATsurfT313K,POPClipid1442perCATsurfT313K,POPClipid1450perCATsurfT313K}.
The surfactants were randomly placed among the lipids to form bilayer structures with
mole fractions of 10\%, 20\%, 30\%, 42\%, or 50\% of surfactant in the POPC bilayer.
All systems contained 50 POPC molecules per leaflet, 6340 TIP3P~\cite{jorgensen83} water molecules and
6, 14, 21, 35, or 50 surfactants per leaflet. 
Chloride counter ions were used
in simulations, because parameters for bromide, the counterion in the experimental data \cite{scherer89},
were not available in the standard Gromacs files for Amber force field.
%\todo{JOE: We cannot present an intricate cool new model for a lipid and claim at the same time that we can't simulate bromide. }
% Yes we can.
The Lipid14 model was used for POPC.
The first 20~ns of the total simulation time of 200~ns
was considered as an equlibration time and was omitted from the analysis.
A reasonable lipid neighbor exchange occured during the simulation.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=7.5cm]{../Fig/POPCstructure.eps}
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.pdf}
  \includegraphics[height=8.4cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
  \caption{\label{simVSexpNOions}
    Top: The chemical structure of POPC and the labeling of the carbon segments.
    Middle: Order parameters of POPC head group, glycerol backbone and acyl chains 
    from simulations with the Lipid14 \cite{dickson14} and the ECC-lipid models
    compared with the experimental values from \cite{ferreira13}.
    The size of the points for the head group order parameters correspond 
    the error bars ($\pm 0.02$ for experiments \cite{botan15,ollila16} and ?? for
    simulations {\bf Add the size of the error bars in simulations.}).
    The size of the points for acyl chains are decreased by a factor of 3 to improve the clarity of the plot.
    Bottom: X-ray scattering form factors from simulations with the Lipid14 \cite{dickson14} and
    the ECC-lipid models compared with experiments~\cite{Kucerka2011}.
  } 
  \todo{Increase the height of the order parameter figure} \\
  \todo{sn-1 and sn-2 order parameters should be somehow labelled in the order parameter figure. Maybe empty and filled points?} \\
  \todo{Would it be possible to increase the size of points for acyl chain order parameters only in z-direction such that
    it would correspond the error bars?} \\
  \todo{x-axis in form factor plot from 0 to 0.6, where experimental data ends. y-axis from 0 to $\sim$2.2 to remove the empty space.} \\
\end{figure}

\begin{table}
  \caption{Area per lipid (APL) values of POPC bilayer without ions from the Lipid14 simulation
    ran in this work and from the literature, from the ECC-lipid model and from experiments.\label{tab:apls} }
  \begin{tabular}{l|c c}
    model          & APL (\AA$^2$)   & Temperature [K] \\
    \hline
    Lipid14                   & 65.1$\pm$ 0.6  &  300 \\
    Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
    \hline
    ECC-lipids                & 62.2$\pm$ 0.6  &  300       \\
    \hline
    experiment \cite{kucerka11} & 64.3  &  303    \\
    \hline
  \end{tabular}
\end{table}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl.pdf}
  \caption{\label{fig:delta_ordPar_CaCl}
    The changes of the head group order parameters of POPC bilayer as a function of CaCl$_2$ concentrations
    are shown from simulations with different force fields together with experimental data 
    (DPPC \cite{akutsu81} and POPC \cite{altenbach84}). 
    Ion concentrations in bulk water are shown in x-axis. 
    Values from simulations are calculated from the of cation number density $C_{np}$
    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
    Simulation data with Lipid14 and \AA{}qvist ion parameters are taken directly from
    Refs.~\cite{lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC}.
  }
  \todo{Maybe experimental points and lines could be larger to make a more clear differentce to simulations} \\
  \todo{Add temperatures for POPC and DPPC experiments} \\
  \todo{y-axis scale of the bottom figure from 59 to 78 to remove the empty space.} \\
  \todo{Empty space from the right column could be reduced.} \\
  \todo{Experimental values from \cite{akutsu81} to be put in the $g_3$ figure:
    the value of $g_3$ order parameter of DPPC was -0.214 
    in the absence of ions [average of two closely spaced splittings] and -0.211 in the presence
    of 0.35 M CaCl$_2$ (at 59 Celcius). No effect of ions could be detected on DPPC bilayers labeled
    at the C-2 segments of both fatty acyl chains.''
  }
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/OrdPars-A-B_L14-ECCL17_q80_sig89_NaCl.pdf}
  \caption{\label{fig:delta_ordPar_NaCl}
    The changes of the head group order parameters of POPC bilayer as a function of NaCl concentrations
    from simulations with different force fields together with experimental data for DPPC~\cite{akutsu81} and POPC~\cite{altenbach84}.
    Ion concentrations in x-axis are calculated as in Fig.~\ref{fig:delta_ordPar_CaCl}.
    Simulation data with Lipid14 and \AA{}qvist ion parameters is taken directly from
    Refs.~\cite{lipid14POPC0mMNaClfiles,lipid14POPC1000mMNaClfiles}.
  }
  \todo{Add also the POPC result to the figure. Ref.\cite{altenbach84} gives for
    POPC at 1M of NaCl $\Delta S^\alpha=(6.1-5.5)*0.00784=0.004704$.} \\
  \todo{Add temperatures to the experimental data.} \\
  \todo{Labeling should be consistent with the previous figure, i.e. experimental data with diamonds etc.} \\
  \todo{x-axis scale from -1 to 1100 to make the point in zero fully visible and to remove empty space.} \\
  \todo{Empty space between figures and from the right column could be reduced.}
\end{figure}

@@

Since significant artefacts are reported in simulations with \AA{}qvist ions is water \cite{??} \todo{Add citations},
we also simulated the Lipid14 model with the ion models by Dang et al.~\cite{smith94,chang1999,dang2006} and
ECC-ions~\cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014} having
more realistic bulk behaviour. Instead of the improvement in the binding behaviour,
we observed overbinding also of Na$^+$ with these ions models, while the results
with Ca$^{2+}$ were barely affected,
%for the CaCl$_2$ interactions with a POPC bilayer described by the Lipid14 model,
as seen in Figs. \ref{fig:delta_ordPar_CaCl} and S??.%\ref{fig:ordPars_waterModels} (in SI), respectively.
\todo{Add OP-response of Lipid14+ECC-ions plot in SI}.
The results are in line with the previous work \cite{catte16},
suggesting that the improvements in the lipid parameters are
required to correctly describe the divalent cation binding to PC lipid bilayers.

%To improve the cation binding behaviour to POPC lipid bilayer we implicitly included the electronic
%polarization to the Lipid14 model by using the ECC approach as described in section \ref{section:ecc}.
The results from the simulations with the ECC-lipid and the ECC-ion models \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}
exhibit an improved behaviour of cation binding to a POPC bilayer in Fig.~\ref{fig:delta_ordPar_CaCl},
showing a good agreement with experiments in the changes of 
the lipid headgroup order parameters as a function of NaCl and CaCl$_2$ concentrations.
Since also the headgroup order parameter
response the to the bound positive charge the ECC-lipid model
was in good agreement with experiments in section \ref{section:boundCHARGE},
we conclude that the model correctly reproduces the binding affinity of
Na$^{+}$ and Ca$^{2+}$ ions to POPC lipid bilayer.
Also the response of the glycerol backbone $g_3$ order parameter to CaCl$_2$ was
signifincantly overestimated in the original Lipid14 model, while the ECC-lipid
gives a good agreement with exepriments, as seen in Fig.~\ref{fig:delta_ordPar_CaCl}.
Also the changes of the P-N vector angle are more prononced
for the Lipid14 model, for which the maximum CaCl$_2$ induced tilting 
toward water phase is approximately 17$^{\circ}$, while the corresponding
value for ECC-lipid model is 6$^{\circ}$.

Furthermore, the overestimated lipid headgroup order parameter changes of POPC in the Lipid14
model as a function of CaCl$_2$ concentration arise from both, the overestimated
binding affinity and sensitivity of the headgroup tilt to the bound positive charge.
This probably applies also to the other lipid models considered in the previous study~\cite{catte16},
emphasizing the importance of the comparison of the lipid headgroup order parameter response
to the bound charge between simulations and experiments as done in section \ref{section:boundCHARGE}.

The ion binding affinity of ECC-lipid model with different water models is compared
in SI. The performance with all water models is better than in in the original Lipid14
parameteres, but the order parameter changes are slightly overestimated
with all four-site water models and with TIP3P.

%\todo{SAMULI: Maybe we should discuss the repeat distances and
%  area per molecules measured at \cite{petrache06,pabst07,uhrikova08}
%  }

@@

\todo{This should be either done properly with the correct $C_I$ or removed. Removing
might be a better idea at this point.}
It was found in the original work~\cite{altenbach84} that 
a ternary complex binding model (i.e.~2~POPC:1~\ce{Ca^{2+}})
provides the best fit to experimental measurements of all considered models in that study. 
In such a model, there is a linear relationship between quantities 
$C_b$, mole fraction of bound \ce{Ca^{2+}} per POPC, and $\sqrt{C_b/C_I}$, 
where $C_I$ is the concentration of free cations at the plane of ion binding~\cite{altenbach84}.
The concentration $C_b$ was obtained from an extrapolation of linear relation 
between deuterium NMR measurements and atomic absorption spectroscopy for low concetrations of \ce{CaCl2}.
Such an extrapolation is valid as long as the mode of \ce{Ca^{2+}} binding 
remains constant throughout the extrapolation range. 
The concentration $C_I$ is determined by using the surface potential by using the Boltzmann equation.
However, Boltzmann theory yields inaccurate results
for divalent cations like \ce{Ca^{2+}}~\cite{Andelman1995}. 
An atomistic simulation, on the other hand, provides these quantities directly without severe assumptions.
Hence we hypothesise that the discrepancy between the results in the experiment~\cite{altenbach84} and 
our simulations likely lays in the fact that 
the assumptions and relations used for determinig concentrations $C_b$ and $C_I$ in the experiment~\cite{altenbach84}
gradually do not hold for higher concentrations of \ce{Ca^{2+}}.

@@

\begin{table}
  \caption{Area per lipid (APL) from different models of POPC with no ions\label{tab:apls_si} }
  \begin{tabular}{l|c c}
    model          & APL (\AA$^2$)   & Temperature [K] \\
    \hline
    Lipid14                   & 65.1$\pm$ 0.6  &  300 \\
    Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
    \hline
    ECC-lipids &        &  \\
    %Lipid14ecc0.80+sigma0.875 &        &  313    \\
    %($4.6\cdot 5.1 \, \mathrm{nm}^2$), 72 lipids patch, OPC3           & 63.2   &   313      \\
    ~ OPC3           & 62.2$\pm$ 0.6   &  300       \\
    ~ OPC3           & 64.2$\pm$ 0.6   &  313       \\
    ~ SPC/E          & 65.1$\pm$ 0.6   &  313       \\
    ~ OPC            & 64.4$\pm$ 0.6   &  313       \\
    ~ TIP4p/2005     & 66.8$\pm$ 0.6   &  313       \\
    %($9.2\cdot 10.2 \, \mathrm{nm}^2$), 288 lipids patch           & 65.5   &  313       \\ %% not done for this model with f_sigma=0.89
    %oMM small patch           & 63.65  &         \\
    %oMM 4xbig patch           & 63.7   &         \\
    \hline
    experiment   & 62.7  &  293    \\
    experiment \cite{kucerka11} & 64.3  &  303    \\
    experiment  & 67.3  &  323    \\
    experiment  & 68.1  &  333    \\
    %experiment POPE  & 56.6 &  303    \\
    \hline
  \end{tabular} \\
  \todo{Result with normal TIP3P missing?}
\end{table}

\begin{table}[btp]
  \caption{Simulation parameters}
  \label{tbl:mdpar}
  \begin{tabular}{ll}
    simulation property & parameter   \\
    \hline
    time-step           & 2~fs         \\
    equilibration time  & 100~ns  \\
    total simulation time     & 300~ns  \\
    temperature         & 313~K       \\
    thermostat          & v-rescale  \cite{bussi07}   \\
    barostat            & Parrinello-Rahman, semi-isotropic \cite{parrinello81} \\
    long-range electrostatics & PME  \cite{darden93}  \\
    cut-off scheme      & Verlet \cite{Pall13}      \\
    Coulomb and VdW cut-off & 1.0~nm \\
    constraints         & LINCS, only hydrogen atoms \cite{hess97} \\
    constraints for water & SETTLE  \cite{miyamoto92} \\
    \hline
  \end{tabular}
\end{table}
